---
title: "Alcohol Analysis"
format: html
date: "Oct 16, 2023"
author: "Catalina Canizares"
editor_options: 
  chunk_output_type: console
---


```{r}
library(tidymodels)
library(tidyverse)
library(dissertationData)
library(gtsummary)
library(mice)

```

```{r}
data("clean_yrbs_2021")

```


```{r}
analysis <-
  clean_yrbs_2021 |>
  select(-c(
    Weight, DrivingDrinking, CurrentlySmokingCigarette,
    StoreSourceVaping, CurrentlySmokingCigar, CurrentlyBingeDrinking,
    UseCondom, NoFruitJuice, NoSalad, NoCarrots, NoSoda, NoBreakfast,
    HIVTested, DentistVisit, NoSportsDrinks, ExcerciseMuscles, Height,
    SexofSexualContact, DrinkingDriver, TextingDriving,
    MoreThan10CigPerDay, CurrentlyUseMarihuana, Sex4OrMorePartners,
    AlcoholOrDrugsSex, BirthControl, WeightLoss, NoFruit,
    NoPotatoes, NoOtherVeggies, NoMilk, PhysicalActivity,
    AttendedPEClass, STDTested, NoDrinksWater, Sunburn,
    EnglishProficiency, CurrentlySmokelessTobacco, MoreThan10Drinks,
    FirstAlcoholBefore13, SexBefore13, SexuallyActive, EverHadSex,
    FirstMarihuanaBefore13, SuicideInjuryMedicalAttention
  ))
```



# Multiple Imputation 

```{r}
md.pattern(analysis$CurrentlyAlcohol)
```

```{r}
ini <- mice(analysis, maxit=0)
meth <- ini$method
meth["CurrentlyAlcohol"] <- "logreg" #logreg' is for binary data

imputed_data <- mice(analysis, method=meth, m=5) # m=5 creates 5 imputed datasets

complete_data_cr <-complete(imputed_data, "long")
saveRDS(complete_data_cr, "imputed_data.rds")


fit <- with(data=imputed_data, exp=glm(CurrentlyAlcohol ~ ParentalMonitoring, family="binomial"))
pooled_results <- pool(fit)
summary(pooled_results)



summary(complete_data_cr)
```


The table presents demographic insights into current alcohol consumption among a group. Out of those surveyed, 27% of females and 19% of males indicated they currently consume alcohol. In terms of racial breakdown, the highest alcohol consumption rates were observed among the American Indian/Alaska Native demographic at 32%, while the lowest was among Asians at 11%. Regarding sexual orientation, bisexual respondents had the highest alcohol consumption rate at 32%. Conversely, heterosexual respondents showed a rate of 22%. When looking at grade levels, 12th graders reported the highest alcohol consumption at 32%, with 9th graders having the lowest rate at 15%.

```{r}
# This options worked to fix the warning: "Stratum (103) has only one
#  PSU at stage 1"
options(survey.adjust.domain.lonely = TRUE)
options(survey.lonely.psu = "adjust")


alcohol_use_demographics <-
  analysis |>
  mutate(CurrentlyAlcohol = case_when(
    CurrentlyAlcohol == 0 ~ "No",
    CurrentlyAlcohol == 1 ~ "Yes"
  )) |>
  mutate(Grade = factor(Grade, levels = c("9", "10", "11", "12"))) |>
  mutate(CurrentlyAlcohol = factor(CurrentlyAlcohol, levels = c("Yes", "No")))


alcohol_use_demographics_srv <- 
survey::svydesign(
  data = alcohol_use_demographics,
  ids = ~psu,
  weights = ~weight,
  strata = ~stratum,
  nest = TRUE
) 

tbl_svysummary(
    data = alcohol_use_demographics_srv,
    include = c(Sex, Race, SexOrientation, Grade),
    sort = c(SexOrientation, Race) ~ "frequency", 
    by = CurrentlyAlcohol, 
    percent = "row") |>
  modify_caption("Demographic Breakdown of Current Alcohol Use: 
                 Based on 2021 YRBS Data") |>
  bold_labels()
```
             

